---
title: "Tidy Verse"
author: "Aaron Zalki"
date: "12/1/2019"
output: html_document
---

# {.tabset}

## Raw Drug-Use Data
### Background on the Data

I am tidying a fivethirtyeight dataset on drug use [Github](https://github.com/fivethirtyeight/data/tree/master/drug-use-by-age). I am transforming the columns of drugs (3 to 28) into rows, in order to perform analysis on the different drugs used by different age groups.
It is important to note that any column header in the dataset with the word 'use' is referring to a percentage of those in an age group who have used the specific drug in the past 12 months. 
Any column header in the dataset with the word 'frequency' is referring to the median number of times a user in an age group used the specific drug in the past 12 months.
For this analysis, I am only interested in the percentage the drug is used among different age groups.


### Loading Data

```{r raw data, include=TRUE}
library(RCurl)
raw_drug <- read.csv(text=getURL("https://raw.githubusercontent.com/fivethirtyeight/data/master/drug-use-by-age/drug-use-by-age.csv"), header=TRUE, stringsAsFactors = FALSE)
head (raw_drug)

```

## Tidy & Transform


I am using the packages tidyr and dplyr to change the raw data to its final form.
```{r transforming, include=TRUE}
library(tidyr)
library(dplyr)
drug_gather <- gather(raw_drug, "Drug", "Percent", 3:28)
drug_gather$Percent <- as.numeric(drug_gather$Percent)
drug_separate <- separate(drug_gather, "Drug", c("Drug", "Usage"))
drug_filter <- filter(drug_separate, Usage=='use')

names (drug_filter) <- c("Age", "Size", "Drug", "Usage", "Percent")

head(drug_filter)
```
## Analysis

### Teenagers

I am using the package ggplot to illustrate the differences found in drug use between teenagers and young adults.

```{r teen, include=TRUE}
library(ggplot2)
teen <- ggplot(subset(drug_filter, Age %in% c("13","14","15", "16", "17", "18","19")), 
aes(x=Drug, y=Percent, fill=Age)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))

teen

```

### Young Adults

```{r young adult, include=TRUE}
young_adults <- ggplot(subset(drug_filter, Age %in% c("18","19","20","21","22-23","24-25")), 
aes(x=Drug, y=Percent, fill=Age)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))


young_adults
```

#### Extended (Harris Dupre)

I will extend the above analyses to more age groups, and compare the frequency of drug use across all age groups.

### Adults 26-49

```{r adult 26-49, include=TRUE}
adults_26_49 <- ggplot(subset(drug_filter, Age %in% c("26-29","30-34","35-49")), 
aes(x=Drug, y=Percent, fill=Age)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))


adults_26_49
```

### Adults 50-65+

```{r adult 49-65+, include=TRUE}
adults_50_65 <- ggplot(subset(drug_filter, Age %in% c("50-64","65+")), 
aes(x=Drug, y=Percent, fill=Age)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))


adults_50_65
```

### Frequency Transform and Analysis Across all Age Groups
#### *Frequency in this context refers to the median frequency of drug usage across an age group for the past 12 months.

```{r, include=TRUE}
drug_filter_freq <- filter(drug_separate, Usage=='frequency')
names (drug_filter_freq) <- c("Age", "Size", "Drug", "Usage", "Frequency")
```

### Teens (Frequency)

```{r teen_freq, include=TRUE}
library(ggplot2)
teen_freq <- ggplot(subset(drug_filter_freq, Age %in% c("13","14","15", "16", "17", "18","19")), 
aes(x=Drug, y=Frequency, fill=Age)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))

teen_freq

```

### Young Adults (Frequency)

```{r young adult freq, include=TRUE}
young_adults_freq <- ggplot(subset(drug_filter_freq, Age %in% c("18","19","20","21","22-23","24-25")), 
aes(x=Drug, y=Frequency, fill=Age)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))


young_adults_freq
```

### Adults 26-49 (Frequency)

```{r adult 26-49 freq, include=TRUE}
adults_26_49_freq <- ggplot(subset(drug_filter_freq, Age %in% c("26-29","30-34","35-49")), 
aes(x=Drug, y=Frequency, fill=Age)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))


adults_26_49_freq
```

### Adults 50-65+ (Frequency)

```{r adult 49-65+ freq, include=TRUE}
adults_50_65_freq <- ggplot(subset(drug_filter_freq, Age %in% c("50-64","65+")), 
aes(x=Drug, y=Frequency, fill=Age)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))


adults_50_65_freq
```
