---
title: "Tidyverse_Foul_Balls"
author: "Harris Dupre"
date: "12/8/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction:

The goal is to create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.

#

We will call the tidyverse package and load in the foul-balls.csv data.

#

```{r}
library("tidyverse")

raw_data <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/foul-balls/foul-balls.csv",na.strings=c("NA", "NULL"))

head(raw_data)
```

#

### Use of tidyverse functions to transform data

#

We will use the tidyverse functions to analyze the exit velocities of foul balls.

#

First, we limit the data to those that contain an exit velocity.Then, using dplyr functions, we'll group the data by the type of hit and summarize the data.

```{r}
my_data <- na.omit(raw_data,cols="exit_velocity")
```
#### Use group_by to select type_of_hit as the analysis group.
```{r}
analysis_df <- group_by(my_data, type_of_hit)
```
#### Use summarize to create summary columns for each type of hit.
```{r}
analysis_df<- summarize(analysis_df, mean=round(mean(exit_velocity),1), median=median(exit_velocity), max = max(exit_velocity), min=min(exit_velocity))
```
#### Make analysis_df into a data frame for plotting.
```{r}
analysis_df <- as.data.frame(analysis_df)

analysis_df
```

#

### Plotting using ggplot functions

#

With a dataframe that contains summary foul ball exit velocity data, we can use the tidyverse package ggplot2 to visualize the summary.

#

### Average exit velocity in ggplot

```{r}
ggplot(analysis_df, aes(x=type_of_hit,y = mean)) +
    geom_bar(width = .75,stat = "identity", position="dodge") +
    ggtitle("Average Exit Velocity of Foul Balls by Type of Hit") +
    labs(x="Type of Hit",y="Average Exit Velocity (mph)") +
    theme(plot.title = element_text(hjust=0.5)) +
    scale_y_continuous(breaks = seq(0,100,by = 5))
```

#

### Max exit velocity in ggplot

```{r}
ggplot(analysis_df, aes(x=type_of_hit,y = max)) +
    geom_bar(width = .75,stat = "identity", position="dodge") +
    ggtitle("Max Exit Velocity of Foul Balls by Type of Hit") +
    labs(x="Type of Hit",y="Max Exit Velocity (mph)") +
    theme(plot.title = element_text(hjust=0.5)) +
    scale_y_continuous(breaks = seq(0,115,by = 5))
```

### Conclusion

#

We can observe that line hits have the fastest average velocity as well as the fastest max velocity for our dataset.

***
\newpage
### Tidyverse Part 2 enhancement - Michael Y.

#### Below is a boxplot which illustrates not only the Max exit velocity of various types of foul balls, but also the median, interquartile range, minimum, and outliers.

```{r boxplot}
ggplot(my_data, aes(x = type_of_hit, y = exit_velocity, fill=type_of_hit)) +
   geom_boxplot() +
   ggtitle("Boxplot: baseball exit velocity")
```